Distributional Regression Forests for Probabilistic Precipitation Forecasting in Complex Terrain

### 2018-10-10 pkg/disttree/inst/draft/rejoinder.tex
some more revision based on reviews
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/partykit@2321 255f4249-2263-4417-b1ea-8498779e27a7

Former-commit-id: db8f0d9ca0385c319b96352e1c0dfe43b43026ff

\documentclass[american,foldmarks=false,noconfig]{uibklttr}
%\documentclass{article}
\usepackage{sfmath,hyperref,color,enumitem}

@@

All changes and additons are explained in much more detail in the
point-to-point reply on the next pages.

@@

The required size of the training data depends heavily on the complexity 
of the model, however, in different applications which we have tested the 
distributional forest on it performed quite well with at least 50 observations. 
Generally, the stopping criteria \texttt{minbucket} and \texttt{minsplit} (both 
determining the minimal number of observations in a node) and \texttt{alpha} 
(the significance level for the statistical tests in the tree algorithm) should 
be kept rather low for a forest model as it is desirable to grow very large trees. 
(FIX ME: reference?)
Therefore, each of the trees in the forest will most likely overfit the subsample 
it is learned on, however, this is evened out by combining them to a forest 
model (with an adequate number of trees). (FIX ME: reference?) 
For that reason, setting the stopping criteria too low does not lead to a decrease 
in goodness of fit, however, the computational costs might increase, especially for 
large data sets. On the other hand, if they are too high, splits might not be detected 
such that the corresponding effects are not captured by the model.

\textit{FIX ME: too long?}

@@

Axams was chosen as the observation station being closest 
to Innsbruck, the capital of Tyrol and the work place of 
three of the authors, but not based on any other particular 
features. Furthermore, it has been checked, whether it 
is an adequate representative of all stations regarding 
the results of our evaluations. This is now also 
confirmed by further analysis in the online supplement which
comes along with our manuscript.

@@

East Tyrol is located south of the main Alpine Ridge and exhibits different
climatological characteristics than the part north of the Alpine Ridge.  It
seems that the choice of covariates and their interactions we made for the
prespecified GAMLSS method is particularly adequate for the region of East
Tyrol for this specific analysis. However, the advantages of the prespecified
GAMLSS models in this particular setting (training: 1985--2008, validation:
2009--2012) even out in a cross-validation framework for most of the stations
in East Tyrol. These results are presented in the online supplement which comes
along with our manuscript showing that the forest performs equally well as the
prespecified GAMLSS and better than the EMOS at most of the stations.

@@

We have rephrased this to 'For each distributional tree a different learning 
data set is obtained ... '

@@

As stopping criteria the minimal number of observations to perform 
a split (\texttt{minsplit}) is set to 50,  the minimal number of 
observations in a terminal node (\texttt{minbucket}) is set to 20 
and the significance level for variable selection (\texttt{alpha}) 
is kept at the default value 1. We have added this information in 
the manuscript in Section~3.2., 'Distributional forest'.

@@

Thank you for pointing this out. We have now rerun the exact 
same evaluations as presented in the paper with different 
distributional assumptions, including a logistic model and 
a hurdle model with a truncated part. The exact results can be 
found in an online supplement.

@@

\begin{review}
3. Is a random forest really the right thing to do? In the 
end, the analyst or meteorologist does not obtain an 
interpretable model and no explanation which variable 
affects the distribution in which direction is possible. 
I think, the reason that focussing on predictions should 
be made clearer and well justified.
\end{review}

@@

\textit{FIX ME: comment on 'there should be a 
better motivation and connection of the methodology and the 
data application'?}

@@

Thank you for your suggestions and the references. 
The prespecification of the distribution is expert knowledge in 
this context as in some of our previous publications we have been 
working on different response distributions and the censored normal 
distribution turned out to be one of the best performing. In this article 
we have not proven whether or not the chosen distribution is the overall
best for this specific application, however, the main part of the article 
is and should be the distributional forest and not the application itself. 
However, we are sure that, given our previous work, the selected methods 
and model assumptions are well performing and well suited for the practical 
example presented.
Moreover, as mentioned before we have tested robustness against 
distributional misspecification. The results show that applying a 
left-censored at zero logistic distribution or a Gaussian hurdle model 
instead of the left-censored at zero Gaussian distribution, as presented 
in the paper, does not lead to any significant improvements in goodness of 
fit of the applied models. The exact results are provided in an online 
supplement which comes with our manuscript.

\textit{FIX ME: refer to Wood, Goicoa, Ugarte}

@@

You are right, there are clearly some stations where other models perform
better than the forest. However, at the majority of stations used in our study
it can at least keep up with them and therefore seems to provide a good and
reasonable alternative that is very easy to apply as no expert knowledge or
additional work in advance is required and the computational effort is kept
relatively small compared to a boosting method. In an online supplement the
same evaluations as for station Axams have been carried out for 14 other
stations, including some of those where the forest does not perform better than
the other models in the setting presented in Figure~8.  It shows that the
differences in performance even out in the cross-validation framework. There,
the forest performs at least equally well compared to the best of the other
models at most stations.

\textit{FIX ME: comments on simulation study? comments on use of
method by meteorologists? (last sentence)}

\begin{reto}
``[\dots] I am wondering whether a meteorologist would really do a
better job on a relative scale of precipitation.'' ist wohl eher eine
rethorische Frage und tut nix zur Sache wenn du mich fragst.
Unkommentiert lassen? Oder Z hat vielleicht eine Idee?
\end{reto}

@@

The folds of the cross-validation are chosen randomly. The data 
set of 28 years is split randomly into 7 blocks of 4 years which 
are not necessarily successive. This is now also explained in the 
description of the cross-validation setup third paragraph of 
Section~3.3..

\begin{review}
8. I found the competing models to be chosen pretty unfair. 
First, for the prespecified GAMLSS, the authors use 
”relevant effects based on expert knowledge”. Why not all? 
Why not making a decision on some information criteria such as AIC/BIC? 
The latter would be done if the analyst decided to use a prespecified 
GAMLSS. I guess, the GAMLSS could improve a lot by this . 
Second, what does a basic EMOS exactly mean? Third, why not to include 
interactions in the boosting algorithm as well? this should be 
conceptually straight forward? Fourth, why are 100 trees chosen 
in the forest? Is this a a default value? May this number not also 
be dependent on the complexity of the model? Why not to choose 
the number of trees similar to what is done in boosting? this 
would also increase the computation time of the forest a lot.
\end{review}

We decided to evaluate two different versions of a GAMLSS: 
the prespecified GAMLSS where the covariates are selected 
in advance based on expert knowledge and the boosted GAMLSS
where all covariates are included. For the latter covariates 
are selected within the boosting algorithm. As stated by 
Hofner et al. (2016) the AIC-based variable selection 
methods implemented in the \textsf{R} package \textbf{gamlss} 
``can be unstable, especially when it comes to selecting 
possibly different sets of variables for multiple 
distribution parameters.''

By a `basic EMOS' we are referring to the linear distributional 
regression model based on the idea presented by Gneiting et al. (2005) 
where only the ensemble mean is included for the location parameter 
and the ensemble standard deviation for the scale parameter.

In the boosted GAMLSS we did not include interactions as this would
lead to an extremely high number of covariates that would be hardly
manageable. Even if we included only the pairwise interactions of all 
80 covariates we would end up with 3160 additional covariates.

% factorial(80)/factorial(80-2)/factorial(2)

\begin{reto}
K\"onnte man eventuell auch im Paper so erw\"ahnen?

\begin{equation}
    \frac{n!}{k! \cdot (n-k)!}~~\mbox{bei}~n=80~\mbox{und}~k=2~\mbox{die}~3160~\mbox{Kombinationen.}
\end{equation}
\end{reto}

By default the number of trees in a distributional forest
is set to 500. However, to keep computation time at a 
reasonable level we decided  to use only 100 trees. 
But of course, the more trees, the better the complexity of 
the model can be captured, especially smooth transitions can 
be modeled better by a higher number of trees.

\textit{FIX ME: comment on 'Why not to choose the number of trees 
similar to what is done in boosting? this would also increase the 
computation time of the forest a lot.'}

@@

@@

@@

The PIT histograms and the residual QQ plots visualize the same
results on different scales. The out-of-sample PIT values are 
either plotted in a histogram or plotted against the corresponding
quantiles of the standard normal distribution. A more detailed 
explanation can be found in Supplement 2 (Stationwise Evaluation) 
which contains both, PIT histograms and residual QQ plots, 
for 15 stations. In the manuscript we decided to show only the 
residual QQ plot as it allows for a better analysis of tail 
behaviour than the PIT histogram.

@@

Figure 6 shows the 10 most important covariates of the forest model
in one particular setting (station Axams, learned on 1985--2008). 
Variable importance has been calculated for all covariates, however,
we decided to present only the top 10 as all following covariates 
are on a rather low and similar level not providing further useful
information. But this top 10 list is not considered for any 
specification of the boosted GAMLSS for which all available 
covariates are included. Out of them, in the exact same setting 
(Axams, 1985--2008), the boosting algorithm selects 28 covariates 
for the location parameter, mainly related to precipitation (7) 
and temperature (11), and 46 covariates for the scale parameter, 
mainly related to precipitation (10), convective available potential 
energy (8) and temperature (15).
% mu: 7 tppow, 1 cape, 2 dswrf, 2 msl, 2 pwat, 3 tcolc, 11 temperature
% sigma: 10 tppow, 8 cape, 1 dswrf, 5 msl, 5 pwat, 2 tcolc, 15 temperature

@@

Applying the stability selection procedure by Meinshausen et al. (2010) and
Hofner et al. (2015) on the same model shows that the 10 covariates with 
maximal selection probability are six covariates based on precipitation and 
one each based on mean sea level pressure, downwards short wave radiation 
flux, temperature and precipitable water respectively.

% load("~/svn/partykit/pkg/disttree/inst/draft/Axams_24to4.rda")
% gb <- Axams_24to4$gb
% sgb <- stabsel(gb, cutoff = 0.6, q = 30)

% R> sgb$selected
% tppow_min.mu         tppow_max.mu    tppow_mean1824.mu   dswrf_mean_mean.mu 
%      3                    4                    7                   25 
% tppow_min.sigma      tppow_max.sigma tppow_sprd2430.sigma  pwat_mean_max.sigma 
%      83                   84                   92                  117 
% t700_mean_max.sigma   msl_diff.sigma 
%      138                  160

% R> head(sort(sgb$max, decreasing = TRUE), 10)
% tppow_max.mu    tppow_mean1824.mu      tppow_max.sigma  t700_mean_max.sigma 
%    1.00                 0.94                 0.91                 0.87 
% dswrf_mean_mean.mu  pwat_mean_max.sigma  tppow_min.sigma tppow_sprd2430.sigma 
%    0.85                 0.84                 0.81                 0.74 
% msl_diff.sigma      tppow_min.mu 
%    0.70                 0.63

% R> head(sort(rowMeans(sgb$phat), decreasing = TRUE), 10)
% tppow_max.mu    tppow_mean1824.mu  t700_mean_max.sigma      tppow_max.sigma 
%    0.99516              0.86018              0.84936              0.84857 
% tppow_min.sigma  pwat_mean_max.sigma   dswrf_mean_mean.mu tppow_sprd2430.sigma 
%    0.79779              0.75813              0.74074              0.65091 
% msl_diff.sigma         tppow_min.mu 
%    0.63928              0.55443

Overall it can be stated that the most important covariates of the boosted 
GAMLSS are similar to the ones of the distributional forest even though they
are not exactly the same ones.

@@

\item[] Ugarte, M. D., Adin, A. and Goicoa, T. (2017). One-dimensional, 
two-dimensional, and three-dimensional B-splines to specify space-time 
interations in bayesian disease mapping: Model fitting and model 
identifiability, \textit{Spatial Statistics} p. to appear.

@@

>>>>

\documentclass[american,foldmarks=false,noconfig]{uibklttr}
%\documentclass{article}
\usepackage{sfmath,hyperref,color,enumitem,soul}

@@

All changes and additions are explained in much more detail in the
point-to-point reply on the next pages.

@@

The required size of the training data depends heavily on the complexity 
of the model, however, in different applications which we have tested the 
distributional forest on it performed quite well with at least 50 observations. \\
Applying the Law of Large Numbers it can be shown that random forests do not overfit, 
not even for an increasing number of trees (Breiman (2001), Hastie et al. (2001), 
Biau et al. (2016)). Therefore, it is recommended to build a large number of trees
as this generally leads to more accurate predictions. However, ``the computational cost 
for inducing a forest increases linearly with the number of trees, so a good choice 
results from a trade-off between computational complexity and accuracy'' 
(Biau et al. (2016)). Following this advice, we decided to build forests consisting 
of 100 trees.\\
Moreover, Hastie et al. (2001) recommend to build trees of maximal depth:
``Our experience is that using full-grown trees seldom costs much, and 
results in one less tuning parameter.''
Thus, the stopping criteria \texttt{minbucket} and \texttt{minsplit} (both 
determining the minimal number of observations in a node) and \texttt{alpha} 
(the significance level for the statistical tests in the tree algorithm) should 
be kept rather low. However, in a distributional framework it also has to be 
considered that the minimal number of observations in a node needs to be high 
enough in order to allow for estimating the distribution parameters reasonably well.
In our application we have set \texttt{minsplit} to 50 and \texttt{minbucket} to 20.
These tuning parameters are now also listed in the main manuscript in Sections~3.2.
and discussed in Section~4..

@@

Axams was chosen as the observation station being closest 
to Innsbruck, the capital of Tyrol and the work place of 
three of the authors, but not based on any other particular 
features. Furthermore, it has been checked, whether it 
is an adequate representative of all stations regarding 
the results of our evaluations. This is now also 
confirmed by further analysis in the online supplement which
comes along with our manuscript and discussed in Section~4..

@@

East Tyrol is located south of the main Alpine Ridge and exhibits different
climatological characteristics than the part north of the Alpine Ridge.  It
seems that the choice of covariates and their interactions we made for the
prespecified GAMLSS method is particularly adequate for the region of East
Tyrol for this specific analysis. However, the advantages of the prespecified
GAMLSS models in this particular setting (training: 1985--2008, validation:
2009--2012) even out in a cross-validation framework for most of the stations
in East Tyrol. These results are presented in the online supplement which comes
along with our manuscript showing that the forest performs equally well as the
prespecified GAMLSS and better than the EMOS at most of the stations.
Moreover, this point is now also discussed in Section~4. in the main manuscript.

@@

We have rephrased this the second paragraph of Section~2.3. which
included this sentence.

@@

As stopping criteria the minimal number of observations to perform 
a split (\texttt{minsplit}) is set to 50,  the minimal number of 
observations in a terminal node (\texttt{minbucket}) is set to 20 
and the significance level for variable selection (\texttt{alpha}) 
is kept at the default value 1. We have added this information in 
the manuscript in Section~3.2., 'Distributional forest' and
also discussed the choice of these values in Section~4. and
in the answer to the second point of the associate editor.

@@

We have applied the left-censored at zero Gaussian distribution
as it has proven to be an adequate choice for this type of 
application in earlier publications.
Moreover, the residual QQ plots and PIT histograms illustrate
that the models are well calibrated employing this distribution.
But thank you very much for the advice to consider a hurdle model!
This is definitely a good idea and therefore we have carried out the
same evaluations as presented in the paper but employing a
Gaussian hurdle model. Additionally, a left-censored at zero logistic
distribution has been applied as well in the same setting in
order to account for heavier tails. The results of these evaluation
can be found in the online supplement 1 (``Different Response Distributions''). 
Overall, it is shown that under the logistic distribution the methods perform 
slightly worse while the hurdle model leads to small improvement in some settings,
however, this comes along with broader variability. The distributional 
forest is proven to be very robust against distributional misspecifications.
Based on these results we decided to keep the distributional assumption
considered in the main manuscript but we have also discussed the tested
alternatives in Section~4..

@@

\begin{review}
3. Is a random forest really the right thing to do? In the 
end, the analyst or meteorologist does not obtain an 
interpretable model and no explanation which variable 
affects the distribution in which direction is possible. 
I think, the reason that focusing on predictions should 
be made clearer and well justified.
\end{review}

@@

We apologize for the confusion the missing equation numbers 
might have caused. They are now added.
We have also rephrased Section~2.3. which now first introduces random 
forests and then distributional forests as an extension of classical 
random forests.

@@

Thank you for your suggestions and the references. These are all
good ideas, however, in our application the number of covariates
is too large in order to keep a model with all interactions manageable.
We have also discussed this in Section~4 of the main manuscript and
rephrased the corresponding sentence in the introduction to ``while 
requiring no meteorological knowledge about the atmospheric processes 
which drive formation of precipitation'' in order to clarify that
this refers only to the variable prespecification in this application.

@@

In order to investigate on these questions we have carried out the full
analysis as presented for station Axams in the main manuscript for 14
further observation stations. It shows that the distributional
forest is very robust against distributional misspecifications
and almost always competitive with the best among the other methods 
or even better. The exact results are provided in 
Supplement 2 (Stationwise Evaluation) and discussed in Section~4. 
Especially for meteorologists intending to apply model output 
statistics (MOS) for several hundreds of observation stations 
(\hl{\texttt{FIXME: reference Reto}})
the distributional forest provides a good and robust alternative. 
Even though computational costs are higher than for an EMOS model, 
they are similar as for a prespecified model and lower than for 
a model applying variable selection based on cross-validation.

@@

The folds of the cross-validation are chosen randomly. The data 
set of 28 years is split randomly into 7 blocks of 4 years which 
are not necessarily successive. This is now also explained in the 
description of the cross-validation setup third paragraph of 
Section~3.3..
We have chosen two different setups: First we learned the model
on the first 24 years and evaluated on the following 4 years
in order to provide a classic time ordered setup. 
However, following (\hl{\texttt{FIXME:} reference Reto})
no influence of time is to be expected. This is also supported by an 
autocorrelation coefficient of -0.007 of the out-of-bag quantile 
residuals of the distributional forest model for station Axams for
a time lag of 1 day.
% ## calculation of autocorrelation:
% load("Axams_24to4.rda") 
% set.seed(4)
% pdf <- predict(Axams_24to4$df, newdata = Axams_24to4$testdata, type = "parameter")
% pit_df <- cbind(0, pnorm(Axams_24to4$testdata[,"robs"], mean = pdf$mu, sd = pdf$sigma))
% pit_df[Axams_24to4$testdata[,"robs"]>0, 1] <- pit_df[Axams_24to4$testdata[,"robs"]>0, 2]
% ac <- acf(qqrplot(pit_df)$residuals)
Then, to evaluate the models in a broader framework a cross-validation 
with the above described choice of folds was applied, for station
Axams in the main manuscript and for 14 further stations in 
Supplement 2 (Stationwise Evaluation).

\begin{review}
8. I found the competing models to be chosen pretty unfair. 
First, for the prespecified GAMLSS, the authors use 
”relevant effects based on expert knowledge”. Why not all? 
Why not making a decision on some information criteria such as AIC/BIC? 
The latter would be done if the analyst decided to use a prespecified 
GAMLSS. I guess, the GAMLSS could improve a lot by this . 
\end{review}

As mentioned in point 5 and also discussed in Section~4, the high
number of covariates (80) restricts the possibilities of variable selection
and specification. For an AIC or BIC based selection an evaluation
of 80 main effects for two parameters would make this a very complex
procedure. Moreover, Hofner et al. (2016) stated that the 
AIC-based variable selection methods implemented in the \textsf{R} 
package \textbf{gamlss} ``can be unstable, especially when it comes 
to selecting possibly different sets of variables for 
multiple distribution parameters.'' For that reason we decided
not to choose such a selection method.

\begin{review}
Second, what does a basic EMOS exactly mean? 
\end{review}

We have adapted the introduction of a basic EMOS in the main manuscript
in Section~ ...
% By a `basic EMOS' we are referring to the linear distributional 
% regression model based on the idea presented by Gneiting et al. (2005) 
% where only the ensemble mean is included for the location parameter 
% and the ensemble standard deviation for the scale parameter.

\begin{review}
Third, why not to include interactions in the boosting algorithm 
as well? this should be conceptually straight forward? 
\end{review}

As now discussed in Section~4 including only all pairwise interactions
of the 80 available variables would already lead to 3160 additional
covariates which would make the model too complex and hardly manageable
regarding computational costs. Therefore, we decided not to include
any interactions in the boosting algorithm.
% factorial(80)/factorial(80-2)/factorial(2)

\begin{review}
Fourth, why are 100 trees chosen in the forest? Is this a a default 
value? May this number not also be dependent on the complexity of 
the model? Why not to choose the number of trees similar to what is 
done in boosting? this would also increase the computation time of 
the forest a lot.
\end{review}

Following Breiman (2001) random forests are very robust and do not overfit, 
not even for a high number of trees. However, computational costs increase 
linearly with the number of trees (Biau et al. (2016)).
Therefore, we followed these recommendations of choosing a high number of
trees, but still not too high in order to keep the computation time at
a reasonable level. Moreover, 100 is a commonly used number of trees that
seems to work very well. Of course, more trees could be applied but 
the performance is usually not significantly worse with 100 trees.
Tuning parameters of forest models, including the number of 
trees (\texttt{ntree}), are now discussed in Section~4 
of the main manuscript.

@@

Thank you for your advice. We have kept only the details regarding
the presented methodology in the appendix of the main manuscript
(Appendix A) and extended Appendix B to an online supplement
(Supplement 2: Stationwise Evaluation).

@@

We used ``EMOS'' as an expression for a standard post-processing
model output statistics (MOS, Gneiting et al. (2005)) with a censored 
distribution. We have revised the introduction of the EMOS model in 
the main manuscript and hope this now explains the model better.

@@

The PIT histograms and the residual QQ plots visualize the same
results on different scales. The out-of-sample PIT values are 
either plotted in a histogram or plotted against the corresponding
quantiles of the standard normal distribution. A more detailed 
explanation can be found in Supplement 2 (Stationwise Evaluation) 
which contains both, PIT histograms and residual QQ plots, 
for 15 stations. In the manuscript we decided to show only the 
residual QQ plot as it allows for a better analysis of tail 
behavior than the PIT histogram.

@@

Figure 6 shows the 10 most important covariates of the forest model
in one particular setting (station Axams, learned on 1985--2008). 
Variable importance has been calculated for all covariates, however,
we decided to present only the top 10 as all following covariates 
are on a rather low and similar level not providing further useful
information. But this top 10 list is not considered for any 
specification of the boosted GAMLSS for which all available 
covariates are included. As for the distributional forest, 
we have also calculated variable importance based on mean increase 
of out-of-bag CRPS for the gamboostLSS model for the exact same 
setting (station Axams, learned on 1985--2008). The resulting 
top 10 variables are listed in the following barplot:

\begin{center}
\includegraphics[width = 0.4\textwidth]{varimp_gb.jpeg}
\end{center}

Overall it can be stated that the most important covariates of the boosted 
GAMLSS are similar to the ones of the distributional forest even though they
are not exactly the same ones and the influence of \texttt{tp\_max} is clearly
stronger in the boosted GAMLSS.

% Out of them, in the exact same setting (Axams, 1985--2008), 
% the boosting algorithm selects 28 covariates 
% for the location parameter, mainly related to precipitation (7) 
% and temperature (11), and 46 covariates for the scale parameter, 
% mainly related to precipitation (10), convective available potential 
% energy (8) and temperature (15).
% mu: 7 tppow, 1 cape, 2 dswrf, 2 msl, 2 pwat, 3 tcolc, 11 temperature
% sigma: 10 tppow, 8 cape, 1 dswrf, 5 msl, 5 pwat, 2 tcolc, 15 temperature

@@

@@

\item[] Ugarte, M. D., Adin, A. and Goicoa, T. (2017). One-dimensional, 
two-dimensional, and three-dimensional B-splines to specify space-time 
interactions in bayesian disease mapping: Model fitting and model 
identifiability, \textit{Spatial Statistics} p. to appear.

@@

\item[] Biau, G. and Scornet, E. (2016).
A Random Forest Guided Tour,
\textit{TEST}
\textbf{25}: 197--227.

\item[] Hastie, T., Tibshirani, R. and Friedman, J. (2001).
The Elements of Statistical Learning,
\textit{Springer New York Inc.}.

\item[] Breiman, L. (2001).
Random Forests,
\textit{Machine Learning}
\textbf{45, 1}: 5--32.
